---
title: "Final Report for Intro to Data Science"
author: "Wyatt Smith - `wsmith4725@floridapoly.edu`"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

# Preparation

Load in the libraries to be used.  Most of the work will be done using functions form the `tidyverse` package.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Import data set
```{r, message=FALSE, warning=FALSE}
platforms <- read_csv("https://github.com/wsmith867/DS_Final_Project/blob/master/data/MoviesOnStreamingPlatforms_updated.csv?raw=true")
```

Variable            | Description                               |
--------------------|-------------------------------------------|
`Title`             | Title name                                | 
`Year`              | Year produced                             | 
`Age`               | Target age group                          |
`IMDb`              | IMDb rating                               |
`Rotten Tomatoes`   | Rotten tomatoes rating                    |
`Netflix`           | Whether the movie is found on Netflix     |
`Hulu`              | Whether the movie is found on Hulu        |
`Prime Video`       | Whether the movie is found on Prime Video |
`Disney+`           | Whether the movie is found on Disney+     |
`Type`              | Movie or TV series                        |
`Directors`         | Directors Name                            | 
`Genre`             | Category of Movie                         | 
`Country`           | Country                                   |
`Language`          | Language available in                     |
`Runtime`           | Duration in minutes                       | 

```{r}
platforms <- na.omit(platforms)
```

