---
title: "Final Report for Intro to Data Science"
author: "Wyatt Smith - `wsmith4725@floridapoly.edu`"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

# Preparation

Load in the libraries to be used.  Most of the work will be done using functions form the `tidyverse` package.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Import data set
```{r, message=FALSE, warning=FALSE}
platforms <- read_csv("https://github.com/wsmith867/DS_Final_Project/blob/master/data/MoviesOnStreamingPlatforms_updated.csv?raw=true")
```


# Introduction
**With the emergence of streaming services, it has become one of, if not, the most popular way to watch movies and/or shows.  At first Netflix was the most recognizable service to use, but now there are multiple choices of platforms to stream from, with each providing its own selection of entertainment.  The aim of this report is to explore and compare the different streaming platforms, from their library size to overall ratings of movies within their libraries.  We will also explore other characteristics of each platform, such as exclusivity of each; given that Netflix has its own line of original movies and shows, and Disney+ would be the main platform for Disney owned movies and shows.  Within the current data set, it classifies whether an entry is a movie or show, but all data entries within this data set are all movies.  Thus, they will be referred to as such.**


# Data Description

```{r}
glimpse(platforms)
```

Variable            | Description                               |
--------------------|-------------------------------------------|
`Title`             | Title name                                | 
`Year`              | Year produced                             | 
`Age`               | Target age group                          |
`IMDb`              | IMDb rating                               |
`Rotten Tomatoes`   | Rotten tomatoes rating                    |
`Netflix`           | Whether the movie is found on Netflix     |
`Hulu`              | Whether the movie is found on Hulu        |
`Prime Video`       | Whether the movie is found on Prime Video |
`Disney+`           | Whether the movie is found on Disney+     |
`Type`              | Movie or TV series                        |
`Directors`         | Directors Name                            | 
`Genre`             | Category of Movie                         | 
`Country`           | Country                                   |
`Language`          | Language available in                     |
`Runtime`           | Duration in minutes                       | 

According to the source of the data set, it was found that Rotten Tomatoes had 69% of its values as NA values, so we shall remove this variable from the data set.  Given that the IMDb variable is still included, we still have a rating variable to use for data analysis.

```{r}
platforms <- platforms %>% 
  select(c(-7))
```

```{r}
platforms <- na.omit(platforms)
```

# Data Exploration

**In terms of just raw numbers, which streaming platform has the largest library?**
```{r}
platforms %>% 
  gather(Netflix, Hulu, `Prime Video`, `Disney+`, key = "platform", value = "value", na.rm = T, convert = T) %>% 
  group_by(platform) %>% 
  filter(value == 1) %>% 
  #count()
  ggplot(aes(x = platform, fill = platform)) +
  geom_bar() +
  labs(title = "Number of movies available on each platform", y = "library size")
```


**Now how many of these are exclusive to each individual platform?**
```{r}
platformNetflix <- platforms %>% 
  filter(Netflix == 1 & Hulu == 0 & `Prime Video` == 0 & `Disney+` == 0) %>% 
  select(ID, Netflix)

platformHulu <- platforms %>% 
  filter(Netflix == 0 & Hulu == 1 & `Prime Video` == 0 & `Disney+` == 0) %>% 
  select(ID, Hulu)

platformPrime <- platforms %>% 
  filter(Netflix == 0 & Hulu == 0 & `Prime Video` == 1 & `Disney+` == 0) %>% 
  select(ID, `Prime Video`)

platformDisney <- platforms %>% 
  filter(Netflix == 0 & Hulu == 0 & `Prime Video` == 0 & `Disney+` == 1) %>% 
  select(ID, `Disney+`)
```

```{r}
platformsExclusive <- data.frame(Netflix = platformNetflix,
                                 Hulu = platformHulu,
                                 Prime = platformPrime,
                                 Disney = platformDisney)
platformsExclusive
```
```{r}
platformsExclusive <- merge(platformNetflix, platformHulu, by = "ID", all = T)
platformsExclusive <- merge(platformsExclusive, platformPrime, by = "ID", all = T)
platformsExclusive <- merge(platformsExclusive, platformDisney, by = "ID", all = T)
platformsExclusive[is.na(platformsExclusive)] <- 0
platformsExclusive
```
```{r}
platformsExclusive %>% 
  gather(Netflix, Hulu, `Prime Video`, `Disney+`, key = "platform", value = "value") %>% 
  filter(value == 1) %>% 
  #group_by(platform) %>% 
  #count()
  ggplot(aes(x = platform, fill = platform)) +
  geom_bar() +
  labs(title = "Movie exclusives on each platform", y = "Num of exclusives", subtitle  = "Exclusive meaning only available on one platform") 
  
```
















